Downward shortwave solar irradiance enhancement by clouds.

### 2024-05-21 article/article.Rmd
Commit 2024-05-21 15:10

<!-- Aerosol change -->
The above clear sky reference values are not enough to describe the long term
variation due to the changes of the atmospheric constituents. As observed
by @Natsis2023, there is long term brightening effect, over Thessaloniki for the
studied period. In order to better describe this change, we used, estimations of AOD
change derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change
of $-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For 2005
to 2023, we derived the trend from the available AERONET data.

To create a unified trend for the long term change of the clear sky irradiance, we
simulated the values of AOD derived from those sources with Libradtran. For both, we
used the AOD at $500\,\text{nm}$, which was inferred by the available Ångström
coefficients, and we choose the SZA of $55^\circ$ as a representative value for all runs.
For the first period, the alpha Ångström coefficient is $1.6$ and the beta
coefficient $0.086$ and $0.059$, for 1997 and 2005 respectively. By interpolating the
resulting clear sky irradiance, we have a trend of $+0.21\,\%$ per year.
For the period 2005 -- 2005 we used the mean monthly values from AERONET, with a
similar simulation scheme. We computed the trend of clear sky radiation for each
month of the year, which gave us a mean clear sky radiation trend of $+0.14\,\%$ per year.

@@

The use of the actual modeled values of clear sky GHI alone, can not provide us with
a robust method to distinguish the CE cases, due to the limited accuracy of the input
data. Our main focus was to positively identify over irradiance events from CEs, thus
we used a relative factor, to create an upper envelope of the clear sky irradiance,
above which, any GHI value can safely attributed to CE.  We evaluated the performance
of the modeled clear radiation for each of atmospheric level of clearness, as
reference, in order to conclude which is the most appropriate.

@@

After evaluating the modeled clear radiation for the different atmospheric
conditions, in relation to the measured GHI data, we choose as a representative of
the clear sky radiation, the case where
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
These values represent a typical atmosphere in Thessaloniki with low load of aerosols
and humidity, which are the main factor that attenuate the GHI, excluding clouds.  To
create the limit of CE identification, we created a two branched threshold, as a
function of SZA. A constant factor for low SZAs, and a higher ramped factor for
higher SZAs (Equation\nobreakspace\ref{eq:CE4}).
This is the criterion of our CE identification.
<!--
\begin{equation}
\text{CE}_\text{Threshold} = \begin{cases}
`r C4_cs_ref_ratio`, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) + `r C4_lowcut_ratio` , & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4a}
\end{equation}
where: $\theta$ solar zenith angle.
-->
<!--
\begin{equation}
\text{CE} : \begin{cases}
 \text{GHI}_\text{i} > `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{i} > \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4b}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modeled clear sky irradiance.
-->
\begin{equation}
\text{CE} : \text{E}_\text{i} > \text{E}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{E}_\text{CSlim,i} = `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{E}_\text{CSlim,i} = \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{E}_\text{CSm,i}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{E}$ the measured
irradiance, $\text{E}_\text{CSm}$ the selected modeled clear sky irradiance, and $i$ each of the one-minute observation.

>>>>

<!-- Aerosol change -->
The above clear sky reference values are not enough to describe the long term
variation, due to the changes of the atmospheric constituents. As observed by
@Natsis2023, there is a long term brightening effect, over Thessaloniki for the studied
period. In order to better describe this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available AERONET data.

To create a unified trend for the long term change of the clear sky irradiance, we
simulated the values of AOD derived from those sources with Libradtran. For both, we
have used the AOD at $500\,\text{nm}$, which was inferred by the available Ångström
coefficients, and we choose the SZA of $55^\circ$ as a representative value for all
runs.

According to @Kazadzis2007, in the period 1997 -- 2005, the mean AOD of
$340\,\text{nm}$ was calculated to $0.403$ with the use of a Brewer
spectrophotometer, with a linear trend of $-3.8±0.93\,\%$ per year.  This corresponds
to a AOD $340\,\text{nm}$ change of $0.0153$ per year.  Using the alpha Ångström
coefficient of $\alpha = 1.6$, this translates to a change in Ångström beta
coefficient of $\beta=0.00272$ per year (or $\beta=0.084$ in 1997 and $\beta=0.059$
in 2005).  By interpolating the resulting clear sky irradiance, we have a trend of
$+0.21\,\%$ per year for the clear sky irradiance.  For the period 2005 -- 2005 we
used the mean monthly values from AERONET, with a similar simulation scheme. We
computed the trend of clear sky radiation for each month of the year, which gave us a
mean clear sky radiation trend of $+0.14\,\%$ per year. We applied this long term
change to the clear sky irradiance, in order to create a more realistic
representation of the irradiance for the whole study period
(Figure\nobreakspace{}\ref{fig:CS-change}).

```{r CS-change, echo=FALSE, fig.cap="Simulated long term change of clear sky irradiance relevant to its climatological values."}
knitr::include_graphics("../images/P_CS_change-1.png")
```

<!-- 1997 0.0839901
<!-- 2005 0.05496013
<!-- # Kazadzis2007 - Nine years of UV aerosol optical depth measurements at Thessaloniki, Greece
<!-- # From Stelios' paper 2007:
<!-- # AOD @340 from Brewer 086
<!-- # Period: 1997-2005
<!-- # The mean AOD for this period was calculated to 0.403 for the Brewer and 0.422 for the CIMEL.
<!-- # A linear regression on the Brewer deseasonalized data reveals a change of -3.8±0.93% per year.
<!-- # From the above data AOD @340 changes by 0.0153 per year (or 0.138 for the 9 years).
<!-- # Using Angstrom a= 1.6  this translates to a change in Angstrom β=0.00272 per year (or β=0.084 in 1997 and β=0.059 in 2005).
-->

@@

The use of the actual modeled values of clear sky GHI, described above , can not provide us with
a robust method to distinguish the CE cases, due to the limited accuracy of the input
data. Our main focus was to positively identify over irradiance events from CEs, 
so we created an upper envelope of the clear sky irradiance,

.............................

above which, any GHI value can safely attributed to CE.  We evaluated the performance
of the modeled clear radiation for each of atmospheric level of clearness, as
reference, in order to conclude which is the most appropriate.

@@

After evaluating the modeled clear radiation for the different atmospheric
conditions, in relation to the measured GHI data, we choose as a representative of
the clear sky radiation, the case where
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
These values represent a typical atmosphere in Thessaloniki with low load of aerosols
and humidity, which are the main factor that attenuate the GHI, excluding clouds.  To
create the limit of CE identification, we created a two branched threshold, as a
function of SZA. A constant factor for low SZAs, and a higher ramped factor for
higher SZAs (Equation\nobreakspace\ref{eq:CE4}).
This is the criterion of our CE identification.
<!--
\begin{equation}
\text{CE}_\text{Threshold} = \begin{cases}
`r C4_cs_ref_ratio`, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) + `r C4_lowcut_ratio` , & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4a}
\end{equation}
where: $\theta$ solar zenith angle.
-->
<!--
\begin{equation}
\text{CE} : \begin{cases}
 \text{GHI}_\text{i} > `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{i} > \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4b}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modelled clear sky irradiance.
-->
<!-- \begin{equation} -->
<!-- \text{CE} : \text{E}_\text{i} > \text{E}_\text{CSlim,i}, \text{where} \begin{cases} -->
<!--  \text{E}_\text{CSlim,i} = `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\ -->
<!-- \text{E}_\text{CSlim,i} = \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{E}_\text{CSm,i}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\ -->
<!-- \text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ -->
<!-- \end{cases}\label{eq:CE4} -->
<!-- \end{equation} -->
<!-- where: $\theta$ is the solar zenith angle, $\text{E}$ the measured -->
<!-- irradiance, $\text{E}_\text{CSm}$ the selected modelled clear sky irradiance, and $i$ each of the one-minute observation. -->

\begin{equation}
\text{CE} : \text{E}_\text{i} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i} \,\,[W/m^2] \label{eq:CE4}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{E}$ the measured
irradiance, $\text{E}_\text{CSm}$ the selected modelled clear sky irradiance, and $i$ each of the one-minute observation.

### 2024-05-21 article/article.tex
Commit 2024-05-21 15:10

The above clear sky reference values are not enough to describe the long
term variation due to the changes of the atmospheric constituents. As
observed by \citet{Natsis2023}, there is long term brightening effect,
over Thessaloniki for the studied period. In order to better describe
this change, we used, estimations of AOD change derived from two
sources. For 1993 -- 2005, \citet{Kazadzis2007} have found a change of
\(-3.8\,\%\) per year at AOD on \(340\,\text{nm}\), using a Brewer
spectrophotometer. For 2005 to 2023, we derived the trend from the
available AERONET data.

To create a unified trend for the long term change of the clear sky
irradiance, we simulated the values of AOD derived from those sources
with Libradtran. For both, we used the AOD at \(500\,\text{nm}\), which
was inferred by the available Ångström coefficients, and we choose the
SZA of \(55^\circ\) as a representative value for all runs.

For the first period, the alpha Ångström coefficient is \(1.6\) and the
beta coefficient \(0.086\) and \(0.059\), for 1997 and 2005
respectively. By interpolating the resulting clear sky irradiance, we
have a trend of \(+0.21\,\%\) per year. For the period 2005 -- 2005 we
used the mean monthly values from AERONET, with a similar simulation
scheme. We computed the trend of clear sky radiation for each month of
the year, which gave us a mean clear sky radiation trend of
\(+0.14\,\%\) per year.

According to \citet{Kazadzis2007}, in the period 1997 -- 2005, the mean
AOD of \(340\,\text{nm}\) was calculated to 0.403 by the Brewer
spectrophotometer, with a linear trend of \(-3.8±0.93\,\%\) per year.
This corresponds to a AOD \(340\,\text{nm}\) change of \(0.0153\) per
year. Using the alpha Ångström coefficient of \(\alpha = 1.6\) this
translates to a change in Ångström beta coefficient of \(\beta=0.00272\)
per year (or \(\beta=0.084\) in 1997 and \(\beta=0.059\) in 2005).

@@

The use of the actual modeled values of clear sky GHI alone, can not
provide us with a robust method to distinguish the CE cases, due to the
limited accuracy of the input data. Our main focus was to positively
identify over irradiance events from CEs, thus we used a relative
factor, to create an upper envelope of the clear sky irradiance, above
which, any GHI value can safely attributed to CE. We evaluated the
performance of the modeled clear radiation for each of atmospheric level
of clearness, as reference, in order to conclude which is the most
appropriate.

@@

After evaluating the modeled clear radiation for the different
atmospheric conditions, in relation to the measured GHI data, we choose
as a representative of the clear sky radiation, the case where
\(\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma\) and
\(w_{h\text{cs}} = w_h - 1\sigma\) (\(\text{GHI}_\text{CSm}\)). These
values represent a typical atmosphere in Thessaloniki with low load of
aerosols and humidity, which are the main factor that attenuate the GHI,
excluding clouds. To create the limit of CE identification, we created a
two branched threshold, as a function of SZA. A constant factor for low
SZAs, and a higher ramped factor for higher SZAs
(Equation\nobreakspace\ref{eq:CE4}). This is the criterion of our CE
identification. \begin{equation}
\text{CE} : \text{E}_\text{i} > \text{E}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{E}_\text{CSlim,i} = 1.04 \cdot \text{E}_\text{CSm,i}, & \text{$\theta \leq 90^\circ$}\\
\text{E}_\text{CSlim,i} = \left ({ 1.04 + \frac{1.04 - 1.04}{90 - 78} \cdot (\theta- 78) } \right ) \cdot \text{E}_\text{CSm,i}, & \text{$ 78^\circ > \theta > 90^\circ$}\\
\text{Excluded measurements}, & \theta > 78^\circ
\end{cases}\label{eq:CE4}
\end{equation} where: \(\theta\) is the solar zenith angle, \(\text{E}\)
the measured irradiance, \(\text{E}_\text{CSm}\) the selected modeled
clear sky irradiance, and \(i\) each of the one-minute observation.

>>>>

The above clear sky reference values are not enough to describe the long
term variation, due to the changes of the atmospheric constituents. As
observed by \citet{Natsis2023}, there is a long term brightening effect,
over Thessaloniki for the studied period. In order to better describe
this change, we used, estimations of AOD change derived from two
sources. For 1993 -- 2005, \citet{Kazadzis2007} have found a change of
\(-3.8\,\%\) per year at AOD on \(340\,\text{nm}\), using a Brewer
spectrophotometer. For 2005 to 2023, we derived the trend from the
available AERONET data.

To create a unified trend for the long term change of the clear sky
irradiance, we simulated the values of AOD derived from those sources
with Libradtran. For both, we have used the AOD at \(500\,\text{nm}\),
which was inferred by the available Ångström coefficients, and we choose
the SZA of \(55^\circ\) as a representative value for all runs.

According to \citet{Kazadzis2007}, in the period 1997 -- 2005, the mean
AOD of \(340\,\text{nm}\) was calculated to \(0.403\) with the use of a
Brewer spectrophotometer, with a linear trend of \(-3.8±0.93\,\%\) per
year. This corresponds to a AOD \(340\,\text{nm}\) change of \(0.0153\)
per year. Using the alpha Ångström coefficient of \(\alpha = 1.6\), this
translates to a change in Ångström beta coefficient of \(\beta=0.00272\)
per year (or \(\beta=0.084\) in 1997 and \(\beta=0.059\) in 2005). By
interpolating the resulting clear sky irradiance, we have a trend of
\(+0.21\,\%\) per year for the clear sky irradiance. For the period 2005
-- 2005 we used the mean monthly values from AERONET, with a similar
simulation scheme. We computed the trend of clear sky radiation for each
month of the year, which gave us a mean clear sky radiation trend of
\(+0.14\,\%\) per year. We applied this long term change to the clear
sky irradiance, in order to create a more realistic representation of
the irradiance for the whole study period
(Figure\nobreakspace{}\ref{fig:CS-change}).

{\centering \includegraphics[width=0.5\linewidth]{../images/P_CS_change-1}

\caption{Simulated long term change of clear sky irradiance relevant to its climatological values.}\label{fig:CS-change}
\end{figure}

@@

The use of the actual modeled values of clear sky GHI, described above ,
can not provide us with a robust method to distinguish the CE cases, due
to the limited accuracy of the input data. Our main focus was to
positively identify over irradiance events from CEs, so we created an
upper envelope of the clear sky irradiance,

\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots\ldots..

above which, any GHI value can safely attributed to CE. We evaluated the
performance of the modeled clear radiation for each of atmospheric level
of clearness, as reference, in order to conclude which is the most
appropriate.

@@

After evaluating the modeled clear radiation for the different
atmospheric conditions, in relation to the measured GHI data, we choose
as a representative of the clear sky radiation, the case where
\(\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma\) and
\(w_{h\text{cs}} = w_h - 1\sigma\) (\(\text{GHI}_\text{CSm}\)). These
values represent a typical atmosphere in Thessaloniki with low load of
aerosols and humidity, which are the main factor that attenuate the GHI,
excluding clouds. To create the limit of CE identification, we created a
two branched threshold, as a function of SZA. A constant factor for low
SZAs, and a higher ramped factor for higher SZAs
(Equation\nobreakspace\ref{eq:CE4}). This is the criterion of our CE
identification.

\begin{equation}
\text{CE} : \text{E}_\text{i} > 15 + 1.04 \cdot \text{E}_\text{CSm,i} \,\,[W/m^2] \label{eq:CE4}
\end{equation} where: \(\theta\) is the solar zenith angle, \(\text{E}\)
the measured irradiance, \(\text{E}_\text{CSm}\) the selected modelled
clear sky irradiance, and \(i\) each of the one-minute observation.
